---
title: "Analysis of Well-Designed Experiment"
author: "u7630977 Tsz Ching Li"
date: "October 2022"
output: pdf_document
---

### Acknowledgment
This project was completed as part of a workshop with provided guidelines and queations. However, **all the code and answers were created by me**.

### Data Unavailability Notice
This project was completed as part of a workshop, where I independently wrote all the code following provided guidelines. Unfortunately, the original dataset used for analysis is no longer available. However, the scripts remain functional and can be adapted for similar datasets.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Temperature and moisture impact strongly on the early stages of a plant’s life cycle. 
Here, we investigate how temperature and moisture affect these early stages of plant development in four Banksia species.

Six polycarbonate-covered steel exclusion shelters were constructed on site to manipulate water availability and assess effects on plant growth as measured by biomass at 6 weeks post-germination. Two garden beds were placed within each shed. Garden beds within a shed were randomised to receive either a warming soil treatment or no warming. Three watering treatments per shed: drought, normal watering, or over-watering were also evaluating within each shed.  

We predicted that reduced moisture and increased temperature would delay plant growth over the 6-week evaluation period.

In our data, we collected Shed number (1-6), Garden Bed (1-2) within each Shed, and watering Section (1-3) within each shed since these factors may contribute to the variation in biomass. 

Import libraries
```{r}
library(tidyverse)
library(grid)
```


Import data. We re-level Water so that it is ordered by amount.
```{r}
banksia <- read_csv("~/Downloads/Banksia data.csv")|>
  mutate(Water = factor(Water, levels = c("low","med","high")))

glimpse(banksia)
```


Summarise the data. How many sheds? How many plants per shed? Create a table to show that the data are balanced.
```{r}
banksia |>
  group_by(Shed) |>
  summarise(number_of_plant = n())
```

__There are six sheds and each sheds is having 24 plants.__

Visualise the layout of temperature treatment across the sheds. 
```{r}
banksia |>
  group_by(Shed, Temperature) |>
  summarise(n = n())

ggplot(banksia, aes(x = as.factor(Bed), y = as.factor(Shed), fill = Temperature)) +
  geom_tile(color = "white") + # color = "white" to outline each section
  scale_fill_brewer(palette = "Set2") + 
  labs(x = "Bed", y = "Shed") +
  facet_grid(~ Section) +
  theme_minimal()
```

Visualise the layout of water treatment across the sheds. 
```{r}
ggplot(banksia, aes(x = as.factor(Bed), y = as.factor(Shed), fill = Water)) +
  geom_tile(color = "white") + # color = "white" to outline each section
  scale_fill_brewer(palette = "Set2") + 
  labs(x = "Bed", y = "Shed") +
  facet_grid(~Section) + # Creating a separate panel for each Shed and Bed
  theme_minimal()

```

Choose a shed and visualise the layout of the species within that shed. 
```{r}
Shed_1 <- banksia |>
  filter(Shed == 1)

ggplot(Shed_1, aes(x = as.factor(Column), y = as.factor(Row), fill = Species)) +
  geom_tile(color = "white") + # color = "white" to outline each section
  scale_fill_brewer(palette = "Set2") + 
  labs(x = "Column", y = "Row") +
  facet_grid(~Section + Bed) + # Creating a separate panel for each Shed and Bed
  theme_minimal() +
  theme(aspect.ratio = 1)
  
```

Exploring biomass and the factors that impact on biomass.
Questions to explore:

```{r}
ggplot(banksia, aes(x = Temperature, y = Biomass, color = Water)) +
  geom_boxplot() +
  facet_wrap(~ Species) +
  theme_minimal()
```

How does temperature affect biomass? Is the temperature effect differ by species? by water treatment? 

__The Temperature have showed a positive correlation with the biomass. Warm temperature has obviously a higher mean, max, min, upper IQ and lower IQ of biomass. All species and humidity have the same pattern.__ 

How does water treatment affect biomass? Does this effect differ by species? by temperature?

__The water treatmet do have a effect to the biomass, but it is not as obvious as the effect of temperature. Also, we can see the effect diverse by species and temperature. Medium water treatment generally have a higher biomass, but Species C dose not show this pattern and Species B in the normal tmperature.__

Fit a linear model to address the research question. Take data structure into account.
```{r}
model <- lm(Biomass ~ Water + Temperature + Species, data = banksia)

summary(model)
```

__The T value of Temperature is the highest, showing the highest effect it have to biomass.__

Interpret the ANOVA table
```{r}
anova(model)
```

Assess the residual plots. Is the model reasonable?
```{r}
library(ggResidpanel) 

resid_panel(model)
```

__The model is reasonable but we can see there is pattern in index plot, some residuals are not lying on the blue line in Q-Q plot, and there are columns higher than the normal distribution curve in histogram, these indicate this suggest non-randomness in the residual, some nuisance factors were not captured in the model.__

Use the summary function and look at the Random Effects. Which factors contribute the most to the variation in biomass?
```{r}
library(lmerTest)

lmer1 <- lmer(Biomass ~ Water + Temperature + Species + 
              (1|Shed) + (1|Plot) + (1|Row) + (1|Column), data = banksia)

summary(lmer1)
```

__For random effect, Shed has a variance of 8.9, which indicate its significant impact on biomass. Temperature in fixed effect has also highly contributed to the variation of biomass, as its estimation is 5.13, which is the highest among fixed effect, there is also a high t value for it, indicating its strong effect.__

As an exercise, try removing one of the random factors from your model. Does it affect the inference?
```{r}
lmer2 <- lmer(Biomass ~ Water + Temperature + Species + (1|Plot) + (1|Row) + (1|Column), data = banksia)

summary(lmer2)
```

__I have removed the Shed random effect, and the residual have increased to 10.81 from 2.771, which means there is more viability left unexplained now. Highlighting Shed have taken a very strong random effect. Also, all the other random effect have become 0 variance after removing Shed, so there are association show between Shed to other factor which is also a known fact when establishing experimental set up__


Use emmeans() to extract the mean biomass and SE estimates from the model for each species, water treatment and temperature. Create a graphic that summarises the results you wish to highlight.
```{r}
library(emmeans)

result <- emmeans(model, ~Species*Water*Temperature)

result
```

```{r}
plot(result, intervals = T)
```

__The plot have shown us there is significant shift to the right for warmer temperature, we can see distinctly two groups (normal temperature and warm temperature). Therefore, we can tell the temperature have obvious impact on the biomass and there are variation among different water treatment and species.__
