---
title: "statistical analysis techniques"
author: "Sharon"
date: "2023-08-23"
output: html_document
---

```{r, echo=FALSE}
library(tidyverse)
library(ggbeeswarm)
library(emmeans)
library(ggResidpanel)
```

Loading data

```{r}
fig_c <-  Kho_2018_Blood_paper_raw_data_Fig1Ci <- read_csv("Data sets for Weeks 5-6 (Lecture 9 amp 10)-20230823/Kho 2018 Blood paper raw data Fig1Ci.csv")

fig_1d <- Kho_2018_Blood_paper_Fig_1Di <- read_csv("~/Downloads/Data sets for Weeks 5-6 (Lecture 9 amp 10)-20230823/Kho 2018 Blood paper Fig 1Di.csv")

```

```{r}
figc_long <- fig_c %>%
  pivot_longer(cols = 1:5, names_to = "status", values_to = "precent")
```


```{r}
ggplot(figc_long, aes(status, precent, color = status)) +
  geom_beeswarm()
```

```{r}
model_figc <- lm(log(precent)~status, data = figc_long)

anova(model_figc)
```

The more samples we have, the better we can get the mean, the SE stand for __Standard Error__ to the mean, it is the uncertainty due to the varaition

type = "response" is the back transform after we have log the data

SE = variation / sqrt(number of sample), the higher the sample number we have, the better we have the mean
```{r}
emmeans(model_figc, ~ status, type = "response")
```

The formula for the t.ratio is log(a) = log(b) = log(a/b)
```{r}
emmeans(model_figc, pairwise~ status, type = "response")$contrast
```

Residual
Visualize whether there is a pattern (we don't want??)
We subtracted the mean
Assumption: is should be normal distribution around the 0
The left bottom is not useful

The histogram: showing r~N(0,σ)
Q-Q plot: showing (e_i, r_i), there is one residual bigger than what we are expecting (the point on right top of the plot)
Residual plot: x = mean, we can see how's the residual distributed, we worry if there is a curve pattern
```{r}
resid_panel(model_figc)
```

```{r}
lm(data = fig_1d, percent ~ status)
```

